Fast gravitational wave parameter estimation without compromises

### 2023-01-23 src/tex/ms.tex
slight title edit
Change package name to texttt

% Title
\title{Fast  gravitational wave parameter estimation with no-compromise}
% Title Options
% \title{Ripples in the Flow: \\ Fast, General Parameter Estimation with Normalizing Flows and Differentiable Waveforms}

@@

\begin{algorithm}
\caption{\textsc{flowMC} pseudocode}\label{alg:cap}
\KwIn{initial position $ip$}
\Parameters{number of training loops $nt$, number of production loops $np$}
\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
\KwResult{$chains$}
$cp \leftarrow ip$\\
\tcc{Training loop}
    \For{$i<nt$}{
        $cc, cp \leftarrow LocalSampling(cp)$\\
        $\Theta \leftarrow TuneNF(cc)$\\
        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $cc \leftarrow Append(cc, chains)$
    }
\tcc{Production loop}
    \For{$i<np$}{
        $c_{local}, cp \leftarrow LocalSampling(cp)$\\
        $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $chains \leftarrow Append(chains, c_{local}, c_{global})$
    }

@@

% GPU helps packing a shit ton of waveform evaluation
Another benefit of accelerators is the parallel evaluation of waveforms. Since
the waveform model we use can be evaluated at any given time or frequency
independently, this means computing a waveform can be trivially parallelized
over frequency bins. Together with the heterodyned likelihood, we can evaluate
the likelihood at $\mathcal{O}(10^7)$ different locations on an \textsc{Nvidia}
A100 GPU.  The high throughput of likelihood evaluations unlocks the potential
of the \textsc{flowMC} sampling algorithms.

@@

% State the distribution of injected population
We sample 1200 events from the distribution of parameters detailed in Table
\ref{tab:parameters}; the same distributions are used as the prior in the PE
process.  We simulate signals over 16 seconds of data, with a minimum frequency
cutoff of 30 Hz and a sampling rate of 2048 Hz.  We draw noise from the design
PSDs for the LIGO Hanford, LIGO Livingston
(\textsc{SimNoisePSDaLIGOZeroDetHighPower}) and Virgo
(\textsc{SimNoisePSDAdvVirgo}) detectors \cite{}.
For both injection and recovery, we make use of the \textsc{IMRPhenomD} waveform
\cite{Khan:2015jqa} via the fully-differentiable implementation presented in the
\textsc{ripple} package \cite{}.

@@

% Brief summary
In this work, we presented a PE pipeline for GW events that is efficient,
flexible and reliable. Our package brings together a number of innovations,
including differentiable waveform models, likelihood heterodyning, and
normalizing-flow enhanced gradient-based sampling. We tested the robustness of
our pipeline, currently built upon \textsc{ripple} and \textsc{flowMC}, on a
set of 1200 synthetic GW events, showing it is robust, unbiased and efficient
enough to handle the large catalogs of detections that will be available in the
near future. We also show that our pipeline can estimate the parameters of
GW150914 and GW170817 in \kw{Quote speed}, demonstrating the potential of our
implementation on real data.

>>>>

% Title
\title{Fast gravitational wave parameter estimation without compromises}
% Title Options
% \title{Ripples in the Flow: \\ Fast, General Parameter Estimation with Normalizing Flows and Differentiable Waveforms}

@@

\begin{algorithm}
\caption{\texttt{flowMC} pseudocode}\label{alg:cap}
\KwIn{initial position $ip$}
\Parameters{number of training loops $nt$, number of production loops $np$}
\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
\KwResult{$chains$}
$cp \leftarrow ip$\\
\tcc{Training loop}
    \For{$i<nt$}{
        $cc, cp \leftarrow LocalSampling(cp)$\\
        $\Theta \leftarrow TuneNF(cc)$\\
        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $cc \leftarrow Append(cc, chains)$
    }
\tcc{Production loop}
    \For{$i<np$}{
        $c_{local}, cp \leftarrow LocalSampling(cp)$\\
        $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $chains \leftarrow Append(chains, c_{local}, c_{global})$
    }

@@

% GPU helps packing a shit ton of waveform evaluation
Another benefit of accelerators is the parallel evaluation of waveforms. Since
the waveform model we use can be evaluated at any given time or frequency
independently, this means computing a waveform can be trivially parallelized
over frequency bins. Together with the heterodyned likelihood, we can evaluate
the likelihood at $\mathcal{O}(10^7)$ different locations on an \textsc{Nvidia}
A100 GPU.  The high throughput of likelihood evaluations unlocks the potential
of the \texttt{flowMC} sampling algorithms.

@@

% State the distribution of injected population
We sample 1200 events from the distribution of parameters detailed in Table
\ref{tab:parameters}; the same distributions are used as the prior in the PE
process.  We simulate signals over 16 seconds of data, with a minimum frequency
cutoff of 30 Hz and a sampling rate of 2048 Hz.  We draw noise from the design
PSDs for the LIGO Hanford, LIGO Livingston
(\textsc{SimNoisePSDaLIGOZeroDetHighPower}) and Virgo
(\textsc{SimNoisePSDAdvVirgo}) detectors \cite{}.
For both injection and recovery, we make use of the \textsc{IMRPhenomD} waveform
\cite{Khan:2015jqa} via the fully-differentiable implementation presented in the
\texttt{ripple} package \cite{}.

@@

% Brief summary
In this work, we presented a PE pipeline for GW events that is efficient,
flexible and reliable. Our package brings together a number of innovations,
including differentiable waveform models, likelihood heterodyning, and
normalizing-flow enhanced gradient-based sampling. We tested the robustness of
our pipeline, currently built upon \texttt{ripple} and \texttt{flowMC}, on a
set of 1200 synthetic GW events, showing it is robust, unbiased and efficient
enough to handle the large catalogs of detections that will be available in the
near future. We also show that our pipeline can estimate the parameters of
GW150914 and GW170817 in \kw{Quote speed}, demonstrating the potential of our
implementation on real data.
